---
title: 'The Winter Olympics: MakeOverMonday 2018 Week7'
author: Aditya Mangal
date: '2018-02-15'
categories:
  - MakeOverMonday
tags:
  - DataAnalysis
  - DataScience
  - MakeOverMonday
  - Olympics
  - Visualizations
  - R
slug: the-winter-olympics-makeovermonday-2018-week7
keywords:
  - tech
summary: Goal of this Visualization task is to create an alternative visualization
  to the [Tableau's visualization for Winter Olympics data for different countries.](https://public.tableau.com/views/TheWinterOlympics/TheWinterOlympics?:embed=y&:showVizHome=no)
readingtime: '5'
coverImage: https://lh3.googleusercontent.com/JsmscZd_PHexzN7qIiIyGKMwqBV0KzMlx5LoQv1p_FE__Me-L20m3JoaJaJeN01JElIYeXdwLcogFKAHPUhFOlXDcjJQpFT7ulAIP7uFgwJ5vczVsLKcxednTuisE3WYDM_V-WGR63huo4HWr9t8-N2_ayXcFAJ5OD4N70XTV983rsxs-PUryxCDmHjFQpGDlkIBF9w_maXIis-HYnxFxJDl7BU2f6h_UOJguug3kwlkkc0DHsTXtG1eQU8z6yuWl1KU1N9Y79MDif2LHE32dk2efds5L6IICOMEbCIX789knem_T6MiKtatUL4GwZvE2yygRQunMlMjHBiKcZibkV6xFwUEty1x5_98gS1yQmg312wLNas6E_CO1F-0uI3dR_UleklXIzJwIAh_4vje1_1TrQsmTcjSk1ioYB1xqg3vQiwI25TaQeth6y64OytrJkliJxJIYM_vXhsk9NAzrSqfpjMb_9mC_R9bE3NuIJpNTKqQ_JShr3Z-u3Zxbc8ySlJttp9ewZ9jbG9jHvib4FEP0y67LK45p-rlwslXCOlDwzV__9rjz-VcN3_Kx0vmh0gaORjiTClg4WO7zjhPBo0kCtzHlJmtmWTHZPPdFlTRKJg8BCDdvGWgmh8m0EiVqekNEUZ7U98Zwalzr8s97Mf8cPYcT3E1=w3270-h1834-no
thumbnailImage: https://lh3.googleusercontent.com/rH18DMGKl5FY_Oh9o6WIoCTt7bDkkwghTNZGDjX6WhqkiUkm8ZWDg5ZvbkQLmNVjPZwZ2ksaKnjAYR2bPiWkLkdhLTN5dLsKfoHNr4Qf6HhDGhVey6eb2hCCiipxs_IUyBewU4QudW7mtj8iNIiYWzp0L_My0ZfgFBAnV1dU5vKVVMm8JEZDzncNX_xuijhjc54OHlvJT7TOZ8Qkh-aG95OPVCWG2vCYN2HaSCC8jUXJ9iYvYam3GmFQ8fQs7kKRLCpRkQgNAvb0STu-5J1KVC3EIrrSIZEkwGCzlQb5E4ZbT2d3mEw8Mo921nq9CG6TvvFMQpKHdXbszn9tBFv25CjnPLwLJ_G1DNc_CbCcbNslhcJSnAvbJKGUPSTu8JHTBpx1NY25I-p5UcA-3pXl50vM4f2Jggcih-Urtr2DXOjpMqWIlwEHcETQ6VbUAQY4aIE71O-WWkJhxekBvIbc28dLE8NP7o7QwnPho6SVdwNuy7milDhI8mL_6FlQ0lDFrEHGdnK03vayfMsv3PWRrsondkCsyHYpW1a5ER9_3MdHTIn57ZdkpNY0eXYICp-3sg_-BSlZSYnWzVdJys7LXMcu4P5zAJXkKQzgEg17Byr_6gyJYDYbb4S1lnoZB4xAKkDxwV_-CWBCc3M6NV1i8MwWEJeS_GCR=w750-h420-no
thumbnailImagePosition: left
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,error = FALSE)
```

# Introduction

Goal of this Visualization task is to create an alternative visualization to the <a href="https://public.tableau.com/views/TheWinterOlympics/TheWinterOlympics?:embed=y&:showVizHome=no" target="_blank">Tableau's visualization for Winter Olympics data for different countries</a>

In this blog post, I'm trying to generate a World Choropleth Map showing the total counts of medals for each country.

# Analysis

## Cleaning up workspace and loading required libraries

```{r Cleanup, results='hide'}
rm(list = ls())
```

```{r Load Libraries, results='hide'}
library(tidyverse) #Data Wrangling
library(httr)
library(readxl) #Data Ingestion
library(ggplot2) #Data Visualization
library(leaflet)
library(rgeos)
library(rgdal)
library(stringr)
```
## Obtaining Data

Reading and viewing the dataset
```{r Reading Data, results='hide'}
GET("https://query.data.world/s/n5nc32oqhtb25hdt3vsa24rd4scs2w", write_disk(tf <- tempfile(fileext = ".xlsx")))
olympics = read_excel(tf)
```
```{r Viewing Data}
olympics
```

Summarizing and getting stats to better understand the dataset

```{r Data Stats}
olympics %>% 
  glimpse()
olympics %>% 
  summary()
```

## Scrubbing data

As per the dataset requirement, East and West Germany are to be grouped with Germany and Soviet Union and the 1992 Unified Team needs to be combined with Russia

```{r Combine countries}
olympics = olympics %>% 
  mutate(Country = recode(Country, "Soviet Union" = "Russia", "Unified Team" = "Russia",
                          "East Germany" = "Germany", "West Germany" = "Germany"))
```

Reading in the ISO-3166 country codes dataset in order to generate the choropleth

```{r Read ISO-3166 country codes dataset, results='hide'}
countryCodes = read_csv("https://raw.githubusercontent.com/lukes/ISO-3166-Countries-with-Regional-Codes/master/all/all.csv")
```
```{r View ISO-3166 country codes dataset}
countryCodes
```

Joining the 2 datasets and verifying if any country name mismatch happening in the 2.

```{r Check if country names mismatch}
olympics %>% 
  left_join(countryCodes, by=c("Country" = "name")) %>% 
  filter(is.na(`alpha-3`)) %>% 
  select(Country) %>% 
  unique()
```

Looks like above 7 countries do not have a corresponding entry in the countryCodes dataset.
Lets try to find out the corresponding names for each of the 7 in the countryCodes dataset.

```{r Find mismatch names in countryCodes}
countryCodes %>% 
  filter(str_detect(str_to_lower(name), "britain")) #United Kingdom of Great Britain and Northern Ireland
countryCodes %>% 
  filter(str_detect(str_to_lower(name), "states")) #United States of America
countryCodes %>% 
  filter(str_detect(str_to_lower(name), "czech")) #Czech Republic
countryCodes %>% 
  filter(str_detect(str_to_lower(name), "russia")) #Russian Federation
countryCodes %>% 
  filter(str_detect(str_to_lower(name), "korea")) #Korea (Democratic People's Republic of) = North Korea, Korea (Republic of) = South Korea
countryCodes %>% 
  filter(str_detect(str_to_lower(name), "yugo")) #Macedonia (the former Yugoslav Republic of)
```

Renaming mismatched countries in olympic dataset based on countryCodes dataset.

```{r Rename mismatches}
olympics = olympics %>% 
  mutate(Country = recode(Country,
                          "Britain" = "United Kingdom of Great Britain and Northern Ireland",
                          "United States" = "United States of America",
                          "Czechoslovakia" = "Czech Republic",
                          "Russia" = "Russian Federation",
                          "North Korea" = "Korea (Democratic People's Republic of)",
                          "South Korea" = "Korea (Republic of)",
                          "Yugoslavia" = "Macedonia (the former Yugoslav Republic of)"))
```

Joining and viewing the 2 datasets

```{r Join Data}
olympics = olympics %>% 
  left_join(countryCodes, by=c("Country" = "name"))
olympics
```

Aggregating per country to find the total number of medals for each country and its corresponding alpha-3 code.

```{r Aggregating per country for Total Medals}
TotalMedalsPerCountry = olympics %>% 
  group_by(Country, `alpha-3`) %>% 
  summarise(TotalMedals = n()) %>% 
  rename(Code = `alpha-3`)
TotalMedalsPerCountry
```

Lets see the top countries based on total number of medals

```{r Top Countries}
TotalMedalsPerCountry %>% 
  arrange(desc(TotalMedals))
```

Germany obtained the most number of medals (377) closely followed by Russia with (341)

## Exploring Data

Lets plot the above data on a map using leaflet.

Loading shape file data set from [World Borders Dataset](http://thematicmapping.org/downloads/world_borders.php). 

```{r}
shape = readOGR("~/Downloads/TM_WORLD_BORDERS_SIMPL-0.3/TM_WORLD_BORDERS_SIMPL-0.3.shp")
names(shape)
```

```{r}
TotalMedalsPerCountry = TotalMedalsPerCountry %>% 
  left_join(tbl_df(shape@data), by = c("Code"="ISO3")) %>% 
  na.omit()

TotalMedalsPerCountry = TotalMedalsPerCountry %>% 
  mutate(label = str_c(sep = " - ", NAME, "TotalMedals", TotalMedals))
```
```{r}
bins = c(0, 10, 20, 30, 50, 100, 150, 200, 300, Inf)
pal = colorBin("RdYlGn", domain = TotalMedalsPerCountry$TotalMedals, bins = bins)

TotalMedalsPerCountry %>% 
  leaflet() %>% 
  addTiles() %>% 
  setView(53.019815, 1.369002, zoom = 1) %>% 
  addCircles(~LON, ~LAT, label = ~label, color = ~pal(TotalMedals), weight = 10)
```


As can be clearly seen above, Norway and USA closely followed in the total medals ranking. 
Visualizing data on a map can provide a clear view of the overall data.

<sub><i>Cover Photo by Willem De Meyer on Unsplash</i></sub>
