---
title: 'India Elections 2019: BJP & Congress Manifestos'' Analysis - Part 2 Economic
  Growth'
author: Aditya Mangal
date: '2019-05-09'
categories:
  - Projects
tags:
  - DataAnalysis
  - DataScience
  - India
  - Python
  - R
  - TextMining
  - Visualizations
slug: india-elections-2019-bjp-congress-manifestos-analysis-part2-economic-growth
keywords:
  - tech
summary: With India's 2019 General Elections around the corner, I thought it'd be
  a good idea to analyse the election manifestos of its 2 biggest political parties,
  BJP and Congress. In this part 2, I'll explore the Economic Growth discussions in
  both manifestos.
readingtime: '10'
draft: no
coverImage: https://lh3.googleusercontent.com/6CZaM8l5GYkmYiKKAQYnu0sbarilrwyz6EfPgWBXjP5O6zIKGlDiwvM95rVhoUWmjoZ_XTcwSBKzML2KvxPQogPSns4I3kgwJvFClsWXMKYbrus-MKfs-WBuGS5L2dTDfPfs3FCWxIg_KsNGzOy3o2L2CvYm5V9febf1sE7zuFAOQnw_opKwPD_eZt9HXngnekY8lUPDBS29RRvHy-owtj1B6nmNwDl2UmtxnOc5SHyr1uvi_vrU7p6sFv408ukkObRysUcT6_jnCvnS_ANLC5CQhg2MULdr-6zclDLRR68xF6UrCeSZv-8O7QFdLySsDhhxDgaHdRi5pH1h0ZoOBFuajcK-7-2nfz6i9tJKnWtXCEZFm_LKOFDNalGmffEesVa_L41I06DbGLQK5Hpl7QCTW6V5FVq8RdF8Kdmy0ozsgnBqQhRkA61JmPXDjnPqWBwiy_3rS0VLo_hsZNj0T2ASsptKcmNgn9900kSj1snl442UTzlQebGwvhiaEdv2En-y3L-Fhe_Fcb0y16rHqf0FGVroQ5MRxkffdjCDt8sLqNMBQ_5g6tgclokYxF0G6RGgNuTt2PzHVK4zvQLp9RQFUReX7I5Vva4R_cOXWCxM0u95rhDiwSAeE-bNyPE5ygiD1OmhaD8IAH2t6uM2X_OHUJZEGrBfaPbQtoZaNwiSQqljzdBpK7bq5Y3lRE3p1ve9iHioa48ZGufs1Aa3Qbkdog=w2748-h1832-no
thumbnailImage: https://lh3.googleusercontent.com/63rXibBvOYO9QbyKunEkNnDmFBAJm_mdO8KtRiWBC6klkT9IvurEugPhRzlFqd0nU8sSUjGkCXLPUqd3g0gJD5CnpxEj4oA_uEg9XrA575XYA05SvUU3zsLGV6jDIi5q7bFjv7br8mTn177YXJgQHrfq7pQqcJ9iOk1QkaSs_rXYTsxEQXVran2HzAavxJ9gYtzCJCM97-ZTXtqVjVCWwY5p4cMwzO_yfogmEJiUJiYc57sAetbM-ntNwksiVO8som6LvJBe7iaAzm21tLBs-z26MtB2epR8sWoHe7RdGw3Yf61vbVHkyxhb36e1CKSzvhj8OQ6oZo_cJcxs3MZKuIZGS8V198G88qfXq0miNpQv2prgqDtgC6cBTQ3fLBfEl2vhYbgmzBZIRMgzaLBgNn1YhAQiWI9pTjEUfD_TQ0bFkmMQQxyc3n5bUphMdlFKR18bGIPnUnXgwqwOwgkX2hJoy-9bm6XcSQ27tUSdL4iQh6-vOo1KV_YVIDhpKbbmBuBT_jH4QPaFspjMq2Dop4qu9m9ytDZMkH4jeN_bYN9JwD-sjovcGPIQFQ4UvkS7JJ6ON24Kmtp9SLCzflCDA9mtMFv2DZx2Myfl0s5eMyuZ3XE0TjgSF67vgAoTD5FmoTrgezkVOSj5diilxesRmbqexIhUUrvMbSRwSjMTpZAh2RiEVQuAqQkUBLAesBCwV4LMyn-HCPfXqf5VF9cXnBzXlA=w750-h500-no
thumbnailImagePosition: left
comments: false
coverCaption: Photo by Annie Spratt on Unsplash
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,error = FALSE)
```


# Introduction

With India's 2019 General Elections around the corner, I thought it'd be a good idea to analyse the election manifestos of its 2 biggest political parties, BJP and Congress. Let's use text mining to understand what each party promises and prioritizes.  
In this part 2, I'll explore the Economic Growth discussions in both manifestos.

# Analysis  

## Load libraries

```{r}
rm(list = ls())
library(tidyverse)
library(pdftools)
library(tidylog)
library(hunspell)
library(tidytext)
library(ggplot2)
library(gridExtra)
library(scales)
library(reticulate)
library(widyr)
library(igraph)
library(ggraph)

theme_set(theme_light())
use_condaenv("stanford-nlp")
```

## Read cleaned data

```{r}
bjp_content <- read_csv("../data/indian_election_2019/bjp_manifesto_clean.csv")
congress_content <- read_csv("../data/indian_election_2019/congress_manifesto_clean.csv")
```


## Economic Growth

This topic is covered in congress' manifesto from Pages 9 to 13 of the pdf and in that of bjp's from pages 13 to 20.

```{r}
bjp_content %>% 
  filter(page >=13,
         page <= 20) -> bjp_content

congress_content %>% 
  filter(page >=9,
         page <= 13) -> congress_content
```

## Most Popular Words

```{r}
plot_most_popular_words <- function(df, 
                                    min_count = 15,
                                    stop_words_list = stop_words) {
  df %>% 
    unnest_tokens(word, text) %>% 
    anti_join(stop_words_list) %>% 
    mutate(word = str_extract(word, "[a-z']+")) %>%
    filter(!is.na(word)) %>% 
    count(word, sort = TRUE) %>% 
    filter(str_length(word) > 1,
           n > min_count) %>% 
    mutate(word = reorder(word, n)) %>%
    ggplot( aes(x=word, y=n)) +
    geom_segment( aes(x=word, xend=word, y=0, yend=n), color="skyblue", size=1) +
    geom_point( color="blue", size=4, alpha=0.6) +
    coord_flip() + 
    theme(panel.grid.minor.y = element_blank(),
          panel.grid.major.y = element_blank(),
          legend.position="none") -> p
  
  return(p)
}
```

```{r}
custom_stop_words <- bind_rows(tibble(word = c("india", "country", "bjp", "congress", "government"), 
                                          lexicon = rep("custom", 5)), 
                               stop_words)
```


```{r fig.width=14}
bjp_content %>% 
  plot_most_popular_words(min_count = 8,
                          stop_words_list = custom_stop_words) + 
  labs(x = "",
       y = "Number of Occurences",
       title = "Most popular words related to Economy Growth in BJP's Manifesto",
       subtitle = "Words occurring more than 8 times",
       caption = "Based on election 2019 manifesto from bjp.org") -> p_bjp

congress_content %>% 
  plot_most_popular_words(min_count = 10,
                          stop_words_list = custom_stop_words) + 
  labs(x = "",
       y = "Number of Occurences",
       title = "Most popular words related to Economy Growth in Congress' Manifesto",
       subtitle = "Words occurring more than 10 times",
       caption = "Based on election 2019 manifesto from inc.in") -> p_congress

grid.arrange(p_bjp, p_congress, ncol = 2, widths = c(10,10))
```

## Common correlated words

```{r}
plot_common_correlated_words <- function(df,
                                         counts_quantile = 0.7,
                                         correlation_threshold = 0.25,
                                         stop_words_list = stop_words) {
  set.seed(123)
  df %>% 
    unnest_tokens(word, text) %>% 
    anti_join(stop_words_list) %>% 
    add_count(word) %>% 
    filter(n > stats::quantile(n, counts_quantile)) %>% 
    pairwise_cor(word, page, sort = TRUE) %>% 
    filter(correlation > correlation_threshold,
         !str_detect(item1, "\\d"),
         !str_detect(item2, "\\d")) %>% 
    graph_from_data_frame() %>%
    ggraph(layout = "fr") +
    geom_edge_link(aes(edge_alpha = correlation), show.legend = FALSE) +
    geom_node_point(color = "lightblue", size = 5) +
    geom_node_text(aes(label = name), repel = TRUE) +
    theme_void() -> p
  
  return(p)
}
```

```{r fig.width=12}
bjp_content %>% 
  plot_common_correlated_words(stop_words_list = custom_stop_words,
                               counts_quantile = 0.85) + 
  labs(x = "",
       y = "",
       title = "Commonly Occurring Correlated Words in BJP's Manifesto",
       subtitle = "Per page correlation higher than 0.25",
       caption = "Based on election 2019 manifesto from bjp.org") -> p_bjp

congress_content %>% 
  plot_common_correlated_words(stop_words_list = custom_stop_words,
                               counts_quantile = 0.85) + 
  labs(x = "",
       y = "",
       title = "Commonly Occurring Correlated Words in Congress's Manifesto",
       subtitle = "Per page correlation higher than 0.25",
       caption = "Based on election 2019 manifesto from inc.in") -> p_congress

grid.arrange(p_bjp, p_congress, ncol = 2, widths = c(12,12))
```

## Basic Search Engine

Lets build a cosine-similarity based simple search engine (instead of the basic keyword-based search that comes with the pdf document), in order to make these documents more easily searchable and gain context using most related lines in the manifestos for a given query. 
Using python's scikit-learn for this.

```{python}
from sklearn.feature_extraction.text import TfidfVectorizer, ENGLISH_STOP_WORDS
from sklearn.metrics.pairwise import linear_kernel

stopwords = ENGLISH_STOP_WORDS

vectorizer_bjp = TfidfVectorizer(analyzer='word', stop_words=stopwords, max_df=0.3, min_df=2)
vector_train_bjp = vectorizer_bjp.fit_transform(r["bjp_content$text"])

vectorizer_congress = TfidfVectorizer(analyzer='word', stop_words=stopwords, max_df=0.3, min_df=2)
vector_train_congress = vectorizer_congress.fit_transform(r["congress_content$text"])

def get_related_lines(query, party="bjp"):
  if (party == "bjp"):
    vectorizer = vectorizer_bjp
    vector_train = vector_train_bjp
  else:
    vectorizer = vectorizer_congress
    vector_train = vector_train_congress
  vector_query = vectorizer.transform([query])
  cosine_sim = linear_kernel(vector_query, vector_train).flatten()
  return cosine_sim.argsort()[:-10:-1]
```

```{r}
get_related_lines <- py_to_r(py$get_related_lines)
```

### Common Popular Words with both BJP & Congress

As we see from the plot above, one of the most popular words in both BJP and Congress' manifesto for economy growth is "farmers". Lets see, what each of them have planned for our farmers. First, BJP.

```{r}
bjp_content %>% 
  slice(get_related_lines("farmer", party = "bjp")) %>% 
  select(text, page, line)
```

This is the excerpt that we find from page 13, as we dug up based on above results - 

> Warehouse Network across the Country - We will build an efficient storage and transport mechanism for agricultural produce.  
 - Our Pradhan Mantri Krishi SAMPADA Yojana highlights our focus on warehousing as a means of increasing farmers’ income. To further expand the warehousing infrastructure in the country, we will establish a National Warehousing Grid along National Highways to ensure necessary logistical linkages for warehousing of agricultural produce.  
 - To enable the farmer to store the agri-produce near his village and sell at a remunerative price at an appropriate time, we will roll out a new Village Storage Scheme of agri-produce. We will provide farmers with loans at cheaper rates on the basis of storage receipt of the agri-produce.

Now, Congress.

```{r}
congress_content %>% 
  slice(get_related_lines("farmer", party = "congress")) %>% 
  select(text, page, line)
```

One of the full excerpts from page 9 related to above results - 

> Congress promises to establish a permanent National Commission on Agricultural Devel- opment and Planning consisting of farmers, agricultural scientists and agricultural economists to examine and advise the government on how to make agriculture viable, competitive and remuner- ative. The recommendations of the Commission shall be ordinarily binding on the government. The Commission will subsume the existing Commission for Agricultural Costs and Prices and recommend appropriate minimum support prices.

### Unique popular words with BJP & Congress

One of the popular words that seems curious from BJP's manifesto is "technology". Let's see what BJP has planned for the use of technology for economic growth.

```{r}
bjp_content %>% 
  slice(get_related_lines("technology", party = "bjp")) %>% 
  select(text, page, line)
```

An excerpt from BJP's manifesto about use of technology as identified from above -  

> We will enable development of young agri-scientists to take advantage of Artificial Intelligence, Machine Learning, Blockchain Technology, Big Data Analytics etc. for more predictive and profitable precision agriculture.

Suprising to see mention of plans for usage of Machine Learning and Blockchain in BJP's manifesto.  

Now, one of the popular words that seems curious from Congress' manifesto is "gst". 

```{r}
congress_content %>% 
  slice(get_related_lines("gst", party = "congress")) %>% 
  select(text, page, line)
```

> MSMEs were badly hit by demonetisation and a flawed GST. Congress promises to devise a rehabilitation plan for MSMEs that were severely affected and help them revive and grow.

Congress is planning for a rehabilitation program for the Micro Small & Medium Enterprises. They also have a plan to redefine MSMEs -  

> MSMEs account for 90 per cent of all employment outside agriculture. The definition of MSMEs based on capital employed is biased against labour. Congress will link the definition of MSME to employment. A business employing 10 persons or less will be ‘micro;’ between 11 and 100 will be ‘small;’ and between 101 and 500 will be ‘medium.’

With all the above analysis, we have developed some idea about the economic growth plans of the 2 parties. In the next post, I'll do a similar analysis for employment and opportunites proposals of them.  

Stay Tuned!

# References  

 - [Part 1 - Data Collection and Cleaning](/2019/05/india-elections-2019-bjp-vs-congress-manifestos-analysis/)  
 - [Part 3 - Employment and Opportunities](/2019/05/india-elections-2019-bjp-congress-manifestos-analysis-part-3-employment-and-opportunities/)
 - For all the parts go to Project Summary Page - [India General Elections 2019 Analysis ](/2019/05/india-general-elections-2019-analysis/)
