---
title: 'Economist''s Visualization Mistakes: TidyTuesday Apr 16 2019'
author: Aditya Mangal
date: '2019-04-16'
categories:
  - TidyTuesday
tags:
  - DataAnalysis
  - TidyTuesday
  - Visualizations
  - R
slug: economist-s-visualization-mistakes
summary: Redrawing Economist's erroneous visualizations as described here - [Mistakes,
  we’ve drawn a few](https://medium.economist.com/mistakes-weve-drawn-a-few-8cdd8a42d368)
readingtime: '10'
draft: no
coverImage: https://lh3.googleusercontent.com/ZQrIa0Kj3JKBAlVNOmHpw5-6eZhyD0cmPqsdKDkMs-FWrNTLdVu5tHL9nH1DZZlRwHl6j8_Mdc7IGHh7H4HB-2CN7fdUxQNjYdJ3_5ETK1jeztm0ZADT2270Nd0nSmib5bZN0AxV8oyODzoYvcHGVDuSwJzlPkbbhk30NdCpte5raPhR4dd8OFOWa9pHvC1LsAMIbt1QJylIrgDytCi1Fain-OiTQ_8LBa2z66lLr-C0p4e2fjv-hJ6sSg7WEW-EFOy1_q6Mo833CVcTD-lER9e6GWUCB8JL6_-gbu_Z9i0A-ZQM-3LtxiB959q7_xn4_yQL_DH0qaSVixdD2asd08GKXm0PSiX5ccQ8_xH1tpt-kitym0FlZgpsdaRmweRso344qiDPGRNOQqWSy3GYwR-mIpslJF-iMJ6B5OwqKFJ644FPml7NS61QGJkNdoKf1DTcn0ARw9uhTW2QTZvRHaM9AOkWrp2mgTrpH6mRMBIWPVv7jpk3SZpZVX5VDt7vts37-hxZgFyWb3FuvV26Kc1Yyz5T9u9UctmwV8Z5vFtld_UIsvOCuqMIUuM_5T_DG4BeVMMY3PeBJ5J7bM6FhR8CxCriRqaHKJ9Xkv-IdxacWdiMVZng7mwDQ9laAAM4anxiEt5uvVydhDp0Fh-5Tk3MBtDG5r8ilYWfSuvw3zQiZgUS-TTfmeULzuqxnTuG3boH4YmqFUG-uSHoTnIalt5oYg=w2602-h1734-no
thumbnailImage: https://lh3.googleusercontent.com/xNXkTy47bwg_XWNQbbqdO5Bk0hoOKmNZua683-YifaJSJaOFN-zsHi5yX-8LkJ4WAZJkhTGCv-kVyYJ8d0RdNXqfDODt3kypwC_nz3lfpNuuzo6m_CjOZhw0AGiHYbvqY8awympo929GydgtY0awlwqIHTtPlqFXj75IzOuR-vpbCtNgK_byginv9D34m7coAQGpdlHDQWUBo2dmaJSDKjvmGn5v4JxViYb6JYYmJDtMAn0SCEmMuz_HN50L_eW9oSXZEDH-Te-4boyPqAZFaor0DmzpxKpWtJsW0ewSqg3Ur6cj9p1VEi_Y9ur_i9fkOvE8WqyspLFG4fHRg2PcLIyAcx3P4l6NrRSiylKYO_1AJ3aXeQdSRmQZ4NRZoD26a2GxOAHi2OXUZScOOFHmC9PwsJ9ESMax8B6R7-CLwUJjvslxoCzYZfRypW7PEZpD-fAKO0hVk6EAyyDkWhFK9QJSHQc3-P6CVYJTzlyVao3iuqM0P7DTtP8WpL7mjYnmdSCV4fbviWZk_NGqnZPNVatM4bXZ1GVeiKlKbEr22HDBqkB0TQVKzWUp0Yp4ZY9XwqUJu_UnIZ8rbEeg88zsxRFGiG-NBkeBzBsThIIsz2Zn2Jv_qpadmxeLHOT5TRkgswOTF1TGTUw1v3v9xYZHTg8S0l4EkZELxNPeTU_kvYjV692_YqIvEn-GC2baEJj5y6kyCR1Od34zIpe6LysY9wDv-w=w750-h500-no
thumbnailImagePosition: left
coverCaption: Photo by Christian Wiediger on Unsplash
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,error = FALSE)
```


# Introduction

From the article on [Mistakes, we’ve drawn a few](https://medium.economist.com/mistakes-weve-drawn-a-few-8cdd8a42d368) - 

"At The Economist, we take data visualisation seriously. Every week we publish around 40 charts across print, the website and our apps. With every single one, we try our best to visualise the numbers accurately and in a way that best supports the story. But sometimes we get it wrong. We can do better in future if we learn from our mistakes — and other people may be able to learn from them, too."

Here I will try and draw the improved plots as suggested by the article on economist or make a version I think is best. 
All this done towards the weekly social data project [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-04-16). 

# Analysis  

## Load libraries

```{r}
rm(list = ls())
library(tidyverse)
library(lubridate)
library(ggplot2)
library(gridExtra)
library(scales)
theme_set(theme_light())
```

## Analyzing Britain's Political Left 

```{r}
corbyn <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-16/corbyn.csv")

corbyn
```


```{r}
corbyn %>% 
  mutate(pct_likes = avg_facebook_likes/sum(avg_facebook_likes)) %>% 
  ggplot(aes(political_group, pct_likes, fill = "red")) + 
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels = percent_format()) + 
  coord_flip() + 
  labs(y = "% of likes over the political groups",
       x = "Political Group",
       title = "Percentage of Average Facebook likes for different political groups",
       caption = "Based on data from The Economist about political left in Britain")
```

## Analyzing decline of dog weights

```{r}
dogs <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-16/dogs.csv")

dogs
```

```{r}
dogs %>% 
  mutate(year = as.factor(year),
         weight_to_neck = avg_weight/avg_neck) %>% 
  ggplot(aes(x = year, y = weight_to_neck)) + 
  geom_line(aes(group = 1)) +
  geom_point() +
  labs(x = "Year",
       y = "Average Weight to Average Neck Ratio",
       title = "Average Weight to Neck Ratio over Years",
       caption = "Based on data from The Economist")
```

## Analyzing Brexit data

```{r}
brexit <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-16/brexit.csv")

brexit
```

```{r}
brexit %>% 
  mutate(date = dmy(date)) %>% 
  ggplot(aes(x = date)) + 
  geom_smooth(aes(y = percent_responding_right, colour = "percent_responding_right"), se = FALSE) + 
  geom_point(aes(y = percent_responding_right, colour = "percent_responding_right")) +
  geom_smooth(aes(y = percent_responding_wrong, colour = "percent_responding_wrong"), se = FALSE) +
  geom_point(aes(y = percent_responding_wrong, colour = "percent_responding_wrong")) + 
  scale_color_manual(labels = c("Right", "Wrong"), values = c("blue", "red")) + 
  labs(x = "Date",
       y = "Response Percentage",
       title = "Response behaviour of people about Brexit over time on the question",
       subtitle = "In hindsight, do you think Britain was right or wrong to vote to leave the EU?",
       caption = "Based on data from The Economist",
       color = "Response")
```

## Analyzing US trade deficit

```{r}
trade <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-16/trade.csv")

trade
```


```{r}
trade %>% 
  mutate(year = as.factor(year),
         trade_deficit = trade_deficit/1000000000,
         manufacture_employment = manufacture_employment/1000000) -> trade

trade %>% 
  ggplot(aes(x = year, y = trade_deficit, fill = "trade_deficit")) + 
  geom_col(show.legend = FALSE) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(x = "Year",
       y = "Trade Deficit in billions",
       title = "US Trade Deficit over years",
       caption = "Based on data from The Economist") -> p1

trade %>% 
  ggplot(aes(x = year, y = manufacture_employment, color = "manufacture_employment")) + 
  geom_line(aes(group = 1), show.legend = FALSE) + 
  geom_point(show.legend = FALSE) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(x = "Year",
       y = "Manufacturing Employment in millions",
       title = "Manufacturing Employment over years",
       caption = "Based on data from The Economist") -> p2

grid.arrange(p1, p2, nrow = 1)
```

## Analyzing pension benefits

```{r}
pensions <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-16/pensions.csv")

pensions
```

```{r}
pensions %>% 
  mutate(spend_per_head = gov_spend_percent_gdp/pop_65_percent) %>% 
  ggplot(aes(x = pop_65_percent, y = gov_spend_percent_gdp, color = country, size = spend_per_head)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(label = country), hjust = -0.15, vjust = 0, show.legend = FALSE) + 
  labs(x = "Percent of population aged 65 or older",
       y = "Percent of government spending on pension benefits as percent of GDP",
       title = "Government Spend vs. Population over 65",
       subtitle = "Size of point represents the spend per head",
       caption = "Based on data from The Economist")
```

## Analyzing EU balance

```{r}
eu_balance <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-16/eu_balance.csv")

eu_balance
```

```{r}
eu_balance %>% 
  mutate(year = as.factor(year),
         account_type = as.factor(account_type),
         country = as.factor(country)) %>% 
  group_by(year, account_type) %>% 
  mutate(perc = value/sum(value)) %>% 
  top_n(5, perc) %>% 
  ungroup() %>% 
  ggplot(aes(x = year, y = value, fill = country)) +
  geom_col() + 
  facet_wrap(~account_type) +
  labs(x = "Year",
       y = "Value in billions of euros",
       title = "Top 5 countries per year and account type",
       subtitle = "Based on percentage of balances",
       caption = "Based on data from The Economist")
```

## Analyzing papers published

```{r}
women_research <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-16/women_research.csv")

women_research
```

```{r fig.width=10}
women_research %>% 
  ggplot(aes(x = field, y = percent_women, color = country, size = percent_women)) + 
  geom_point() + 
  scale_size(guide = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Field of study",
       y = "Percentage of Women",
       title = "Women among researchers with papers published 2011-15 as % of total by field of study for 12 countries",
       subtitle = "Size of point also represents percentage of women",
       color = "Country",
       caption = "Based on data from The Economist")
```
