---
title: 'India Elections 2019: BJP & Congress Manifestos'' Analysis - Part 7 Education
  Healthcare & Miscellaneous'
author: Aditya Mangal
date: '2019-05-13'
categories:
  - Projects
tags:
  - DataAnalysis
  - DataScience
  - India
  - Python
  - R
  - SentimentAnalysis
  - TextMining
  - Visualizations
slug: india-elections-2019-bjp-congress-manifestos-analysis-part-7-education-healthcare-misc
keywords:
  - tech
summary: With India's 2019 General Elections around the corner, I thought it'd be
  a good idea to analyse the election manifestos of its 2 biggest political parties,
  BJP and Congress. In this part 7, I'll explore the Education, Health Care and other
  miscellaneous discussions in both manifestos.
readingtime: '10'
draft: no
coverImage: https://lh3.googleusercontent.com/6CZaM8l5GYkmYiKKAQYnu0sbarilrwyz6EfPgWBXjP5O6zIKGlDiwvM95rVhoUWmjoZ_XTcwSBKzML2KvxPQogPSns4I3kgwJvFClsWXMKYbrus-MKfs-WBuGS5L2dTDfPfs3FCWxIg_KsNGzOy3o2L2CvYm5V9febf1sE7zuFAOQnw_opKwPD_eZt9HXngnekY8lUPDBS29RRvHy-owtj1B6nmNwDl2UmtxnOc5SHyr1uvi_vrU7p6sFv408ukkObRysUcT6_jnCvnS_ANLC5CQhg2MULdr-6zclDLRR68xF6UrCeSZv-8O7QFdLySsDhhxDgaHdRi5pH1h0ZoOBFuajcK-7-2nfz6i9tJKnWtXCEZFm_LKOFDNalGmffEesVa_L41I06DbGLQK5Hpl7QCTW6V5FVq8RdF8Kdmy0ozsgnBqQhRkA61JmPXDjnPqWBwiy_3rS0VLo_hsZNj0T2ASsptKcmNgn9900kSj1snl442UTzlQebGwvhiaEdv2En-y3L-Fhe_Fcb0y16rHqf0FGVroQ5MRxkffdjCDt8sLqNMBQ_5g6tgclokYxF0G6RGgNuTt2PzHVK4zvQLp9RQFUReX7I5Vva4R_cOXWCxM0u95rhDiwSAeE-bNyPE5ygiD1OmhaD8IAH2t6uM2X_OHUJZEGrBfaPbQtoZaNwiSQqljzdBpK7bq5Y3lRE3p1ve9iHioa48ZGufs1Aa3Qbkdog=w2748-h1832-no
thumbnailImage: https://lh3.googleusercontent.com/63rXibBvOYO9QbyKunEkNnDmFBAJm_mdO8KtRiWBC6klkT9IvurEugPhRzlFqd0nU8sSUjGkCXLPUqd3g0gJD5CnpxEj4oA_uEg9XrA575XYA05SvUU3zsLGV6jDIi5q7bFjv7br8mTn177YXJgQHrfq7pQqcJ9iOk1QkaSs_rXYTsxEQXVran2HzAavxJ9gYtzCJCM97-ZTXtqVjVCWwY5p4cMwzO_yfogmEJiUJiYc57sAetbM-ntNwksiVO8som6LvJBe7iaAzm21tLBs-z26MtB2epR8sWoHe7RdGw3Yf61vbVHkyxhb36e1CKSzvhj8OQ6oZo_cJcxs3MZKuIZGS8V198G88qfXq0miNpQv2prgqDtgC6cBTQ3fLBfEl2vhYbgmzBZIRMgzaLBgNn1YhAQiWI9pTjEUfD_TQ0bFkmMQQxyc3n5bUphMdlFKR18bGIPnUnXgwqwOwgkX2hJoy-9bm6XcSQ27tUSdL4iQh6-vOo1KV_YVIDhpKbbmBuBT_jH4QPaFspjMq2Dop4qu9m9ytDZMkH4jeN_bYN9JwD-sjovcGPIQFQ4UvkS7JJ6ON24Kmtp9SLCzflCDA9mtMFv2DZx2Myfl0s5eMyuZ3XE0TjgSF67vgAoTD5FmoTrgezkVOSj5diilxesRmbqexIhUUrvMbSRwSjMTpZAh2RiEVQuAqQkUBLAesBCwV4LMyn-HCPfXqf5VF9cXnBzXlA=w750-h500-no
thumbnailImagePosition: left
comments: yes
coverCaption: Photo by Annie Spratt on Unsplash
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,error = FALSE)
```


# Introduction

With India's 2019 General Elections around the corner, I thought it'd be a good idea to analyse the election manifestos of its 2 biggest political parties, BJP and Congress. Let's use text mining to understand what each party promises and prioritizes.  
In this part 7, I'll explore the Education, Health Care and other miscellaneous discussions in both manifestos.

# Analysis  

## Load libraries

```{r}
rm(list = ls())
library(tidyverse)
library(pdftools)
library(tidylog)
library(hunspell)
library(tidytext)
library(ggplot2)
library(gridExtra)
library(scales)
library(reticulate)
library(widyr)
library(igraph)
library(ggraph)

theme_set(theme_light())
use_python("~/anaconda3/bin/python")
```

## Read cleaned data

```{r}
bjp_content <- read_csv("../data/indian_election_2019/bjp_manifesto_clean.csv")
congress_content <- read_csv("../data/indian_election_2019/congress_manifesto_clean.csv")
```


## Education, Health Care and Miscellaneous

This topic is covered congress' manifesto from Pages 24 to 27 of the pdf and in that of bjp's from pages 23 and 29 to 30 and 36 to 37. 

```{r}
bjp_content %>% 
  filter(between(page, 23, 23) | between(page, 29, 30) | between(page, 36, 37)) -> bjp_content

congress_content %>% 
  filter(between(page, 24, 27)) -> congress_content
```

## Most Popular Words

```{r}
plot_most_popular_words <- function(df, 
                                    min_count = 15,
                                    stop_words_list = stop_words) {
  df %>% 
    unnest_tokens(word, text) %>% 
    anti_join(stop_words_list) %>% 
    mutate(word = str_extract(word, "[a-z']+")) %>%
    filter(!is.na(word)) %>% 
    count(word, sort = TRUE) %>% 
    filter(str_length(word) > 1,
           n > min_count) %>% 
    mutate(word = reorder(word, n)) %>%
    ggplot( aes(x=word, y=n)) +
    geom_segment( aes(x=word, xend=word, y=0, yend=n), color="skyblue", size=1) +
    geom_point( color="blue", size=4, alpha=0.6) +
    coord_flip() + 
    theme(panel.grid.minor.y = element_blank(),
          panel.grid.major.y = element_blank(),
          legend.position="none") -> p
  
  return(p)
}
```

```{r}
custom_stop_words <- bind_rows(tibble(word = c("india", "country", "bjp", "congress", "government"), 
                                          lexicon = rep("custom", 5)), 
                               stop_words)
```


```{r fig.width=14}
bjp_content %>% 
  plot_most_popular_words(min_count = 7,
                          stop_words_list = custom_stop_words) + 
  labs(x = "",
       y = "Number of Occurences",
       title = "Most popular words related to Education, Healthcare & Misc in BJP's Manifesto",
       subtitle = "Words occurring more than 7 times",
       caption = "Based on election 2019 manifesto from bjp.org") -> p_bjp

congress_content %>% 
  plot_most_popular_words(min_count = 12,
                          stop_words_list = custom_stop_words) + 
  labs(x = "",
       y = "Number of Occurences",
       title = "Most popular words related to Education, Healthcare & Misc in Congress' Manifesto",
       subtitle = "Words occurring more than 12 times",
       caption = "Based on election 2019 manifesto from inc.in") -> p_congress

grid.arrange(p_bjp, p_congress, ncol = 2, widths = c(10,10))
```

## Common correlated words

```{r}
plot_common_correlated_words <- function(df,
                                         counts_quantile = 0.7,
                                         correlation_threshold = 0.25,
                                         stop_words_list = stop_words) {
  set.seed(123)
  df %>% 
    unnest_tokens(word, text) %>% 
    anti_join(stop_words_list) %>% 
    add_count(word) %>% 
    filter(n > stats::quantile(n, counts_quantile)) %>% 
    pairwise_cor(word, page, sort = TRUE) %>% 
    filter(correlation > correlation_threshold,
         !str_detect(item1, "\\d"),
         !str_detect(item2, "\\d")) %>% 
    graph_from_data_frame() %>%
    ggraph(layout = "fr") +
    geom_edge_link(aes(edge_alpha = correlation), show.legend = FALSE) +
    geom_node_point(color = "lightblue", size = 5) +
    geom_node_text(aes(label = name), repel = TRUE) +
    theme_void() -> p
  
  return(p)
}
```

```{r fig.width=12}
bjp_content %>% 
  plot_common_correlated_words(stop_words_list = custom_stop_words,
                               counts_quantile = 0.85) + 
  labs(x = "",
       y = "",
       title = "Commonly Occurring Correlated Words in BJP's Manifesto",
       subtitle = "Per page correlation higher than 0.25",
       caption = "Based on election 2019 manifesto from bjp.org") -> p_bjp

congress_content %>% 
  plot_common_correlated_words(stop_words_list = custom_stop_words,
                               counts_quantile = 0.85) + 
  labs(x = "",
       y = "",
       title = "Commonly Occurring Correlated Words in Congress's Manifesto",
       subtitle = "Per page correlation higher than 0.25",
       caption = "Based on election 2019 manifesto from inc.in") -> p_congress

grid.arrange(p_bjp, p_congress, ncol = 2, widths = c(12,12))
```

## Basic Search Engine

Lets build a cosine-similarity based simple search engine (instead of the basic keyword-based search that comes with the pdf document), in order to make these documents more easily searchable and gain context using most related lines in the manifestos for a given query. 
Using python's scikit-learn for this.

```{python}
from sklearn.feature_extraction.text import TfidfVectorizer, ENGLISH_STOP_WORDS
from sklearn.metrics.pairwise import linear_kernel

stopwords = ENGLISH_STOP_WORDS

vectorizer_bjp = TfidfVectorizer(analyzer='word', stop_words=stopwords, max_df=0.3, min_df=2)
vector_train_bjp = vectorizer_bjp.fit_transform(r["bjp_content$text"])

vectorizer_congress = TfidfVectorizer(analyzer='word', stop_words=stopwords, max_df=0.3, min_df=2)
vector_train_congress = vectorizer_congress.fit_transform(r["congress_content$text"])

def get_related_lines(query, party="bjp"):
  if (party == "bjp"):
    vectorizer = vectorizer_bjp
    vector_train = vector_train_bjp
  else:
    vectorizer = vectorizer_congress
    vector_train = vector_train_congress
  vector_query = vectorizer.transform([query])
  cosine_sim = linear_kernel(vector_query, vector_train).flatten()
  return cosine_sim.argsort()[:-10:-1]
```

```{r}
get_related_lines <- py_to_r(py$get_related_lines)
```

### Common Popular Words with both BJP & Congress

As we see from the plot above, one of the most popular words in both BJP and Congress' manifesto is "children". Lets see, what each of them have planned for the children of our country. First, BJP.

```{r}
bjp_content %>% 
  slice(get_related_lines("children", party = "bjp")) %>% 
  select(text, page, line)
```

This is the excerpt that we find from page 29, as we dug up based on above results - 

> We consider it our duty to provide a conducive environment for talented children in our country to facilitate them to unleash their talent. For this, we will initiate a ‘Prime Minister Innovative Learning Program’ to bring together such children for a certain period in a year in one place from all over the country and provide them with facilities and resources for them to excel. This would instill a culture of lateral thinking and innovation among them.

Now, Congress.

```{r}
congress_content %>% 
  slice(get_related_lines("children", party = "congress")) %>% 
  select(text, page, line)
```

One of the excerpts from page 25 related to above results - 

> Congress promises that school education from Class I to Class XII in public schools shall be compulsory and free. Suitable amendments will be made in the Right to Education Act, 2009. We will end the practice of charging special fees for different purposes in public schools.

### Unique popular words with BJP & Congress

One of the popular words that seems curious from BJP's manifesto is "yoga". 

```{r}
bjp_content %>% 
  slice(get_related_lines("yoga", party = "bjp")) %>% 
  select(text, page, line)
```

An excerpt from BJP's manifesto about use of yoga as identified from above -  

> We will further continue our efforts to promote Yoga globally as the world celebrates 21st June as the International Yoga Day. We will promote Yoga as a vital method to achieve physical wellness and spiritual rejuvenation across the globe and will continue to work towards training of Yoga practitioners. We will undertake a rapid expansion of Yoga health hubs, Yoga tourism and research in Yoga.

Now, one of the popular words that seems curious from Congress' manifesto is "sports". 

```{r}
congress_content %>% 
  slice(get_related_lines("sports", party = "congress")) %>% 
  select(text, page, line)
```

> Congress will take suitable measures to ensure that the Constitution of each sports body is in accordance with certain norms and principles, that elections are held according to them, and that there is adequate representation in each sports body for active players, former players and women.

With all the above analysis, we have developed some idea about the Education, Healthcare and other miscellaneous plans of the 2 parties.  


With this I conclude my analysis of the manifestos of BJP and Congress for 2019 general elections. I highly encourage all my readers to make their choice carefully in these pivotal elections and even though these analyses give you a high level idea about each party, maybe try and go through the entire manifestos to make a much informed decision about your choice and ensure accountability from our governments.  

Let's hope for and build a better India. All the best!

# References  

 - [Part 6 - National Security](/2019/05/india-elections-2019-bjp-congress-manifestos-analysis-part-6-national-security/)  
 - For all the parts go to Project Summary Page - [India General Elections 2019 Analysis ](/2019/05/india-general-elections-2019-analysis/)
